---
title: "Ethics Check"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ethics Approval at Scale}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r}
#| label: setup
#| message: false
devtools::load_all(".")
library(dplyr)
```

This vignette shows how to run the `ethics_check` module, first on a single paper and then on the built-in `psychsci` corpus of 250 open-access papers from Psychological Science.

## Running `ethics_check`

`ethics_check` searches each paper for ethics approval statements and for signals of live data collection from human or animal participants. It returns a standard module list with a `summary_table` that records, per paper:

- `ethics_approved` — whether an ethics statement was found
- `ethics_statements` — the matching ethics-related sentence(s), if any
- `needs_ethics` — whether live data collection was detected (so ethics approval would be expected)
- `live_data_statements` — the matching live-data-collection sentence(s), if any

```{r}
paper <- psychsci$`0956797620955209`
result <- module_run(paper, "ethics_check")

result$summary_table
```

The `report` explains the verdict by quoting the relevant sentence(s) — the live-data sentence that triggered the expectation of an ethics statement, and the ethics statement itself if one was found.

```{r, results = 'asis'}
cat(result$report)
```

## Generating a report

You can also generate a standalone report for a paper that includes only the `ethics_check` module, by passing it as the only entry in the `modules` argument of `report()`.

```{r, eval = FALSE}
paper2 <- psychsci$`0956797614522816`
report(paper2, modules = "ethics_check")
```

This writes an HTML file (by default named after the paper's `paper_id`) to the working directory; the saved path is available as `attr(result, "save_path")`. Use `output_format = "qmd"` if you want the editable Quarto source instead of rendered HTML.

## Running on the `psychsci` corpus

```{r}
result_ps <- module_run(psychsci, "ethics_check")
```

The `summary_table` has one row per paper.

```{r}
result_ps$summary_table |> head()
```

The traffic light gives an overall verdict for the corpus, and `summary_text` a human-readable count.

```{r}
result_ps$traffic_light
result_ps$summary_text
```

## Examining papers without ethics statements

To inspect individual papers that appear to involve live data collection but lack an ethics approval statement:

```{r}
missing_ps <- result_ps$summary_table |>
  filter(needs_ethics & !ethics_approved)

nrow(missing_ps)
missing_ps$paper_id
```

The `live_data_statements` column shows the sentence(s) that triggered the expectation of an ethics statement for each of these papers.

```{r}
missing_ps$live_data_statements[1:3]
```

You can also look at the detected sentences for any of these papers by filtering the full `table` output.

```{r}
result_ps$table |>
  filter(paper_id == missing_ps$paper_id[1])
```

## Validation against other corpora

The patterns in `ethics_check` were validated against two corpora that are not bundled with the package: 749 papers from *Collabra: Psychology* and 855 papers from the *Judgment and Decision Making* journal. These corpora were chosen because they sit outside the Psychological Science distribution the patterns were originally developed on, so they provide an independent check on generalization.

Validation proceeded in three rounds:

- **Round 1.** `ethics_check` was run on both full corpora, and a stratified sample of 106 papers (57 Collabra, 49 JDM) was drawn across the four cells of `ethics_approved` x `needs_ethics`. Each sampled paper's full text was read and independently labeled with a ground-truth `manual_ethics_approved` and `manual_needs_ethics` verdict plus a free-text justification, in batches of about 10 papers at a time. Comparing module output to these manual labels gave a combined "missing ethics" flag (`needs_ethics & !ethics_approved`) accuracy of 80.2% (7 false positives, 14 false negatives out of 106). The largest error source was the short committee abbreviations REC/REB/DEC colliding with unrelated jargon (e.g. "REC" as a recognition-heuristic model name, "DEC" as a decision-task label), which inflated `ethics_approved` false positives.
- **Fix.** The REC/REB/DEC/ERB patterns were constrained to only match when an ethics/approval/committee/board/panel/institutional/protocol word appears within about 30 characters in the same sentence (IRB and IEC were left unconstrained, since they were not observed to collide in practice).
- **Round 2.** A fresh stratified sample of 100 papers (43 Collabra, 57 JDM), excluding all round-1 papers, was drawn and manually labeled the same way. Combined-flag accuracy rose to 91% (8 false positives, 1 false negative out of 100); `ethics_approved` accuracy was 96% and `needs_ethics` accuracy was 93%. Remaining disagreements were genuinely ambiguous cases (e.g. secondary analyses of other researchers' data, methods-guidance articles instructing future authors to obtain their own approval, and one borderline case with no formal "approved by" sentence but compliance language plus informed consent).
- **Round 3.** With the fixes in place, `ethics_check` was re-run on the full corpora to confirm the sample-level improvements held at scale: 420 of 749 Collabra papers and 30 of 855 JDM papers were flagged `ethics_approved`, with 649 and 776 papers respectively flagged `needs_ethics`.

This validation work (scripts, sampled data, and manual annotations) lives in `_stuff/ethics_sample/` in the package source repository and is not part of the installed package.
