---
title: "Modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(papercheck)
library(dplyr)
```

Papercheck is designed modularly, so you can add modules to check for anything. It comes with a set of pre-defined modules, and we hope people will share more modules.

## Module List

You can see the list of built-in modules with the function below.

```{r, results='asis'}
module_list()
```

## Built-in Modules

Below, we will demonstrate the use of each built-in module, first on a single paper and then a list of papers, the `psychsci` list of 250 open-access papers from Psychological Science.

```{r}
paper <- psychsci$`0956797620955209`
```


### all_p_values

List all p-values in the text, returning the matched text (e.g., 'p = 0.04') and document location in a table.

```{r}
all_p <- module_run(paper, "all_p_values")

all_p$table # print table
```


If you run this module on all 250 papers, you will get more rows than you probably want to print in the full table one row for every p-value in each paper), so you can print the summary table, which gives you one row per paper.

```{r}
all_p_ps <- module_run(psychsci, "all_p_values")

all_p_ps$summary
```

You can still access the full table for further processing.

```{r}
all_p_ps$table |>
  count(text, sort = TRUE) |>
  head()
```


### all_urls

List all the URLs in the main text. There will, of course, be a few false positives when text in the paper is formatted as a valid URL. 

```{r}
all_urls <- module_run(paper, "all_urls")

all_urls$table
```


```{r}
all_urls_ps <- module_run(psychsci, "all_urls")

all_urls_ps$summary
```

### imprecise_p

List any p-values that may have been reported with insufficient precision (e.g., p < .05 or p = n.s.). 

```{r}
imprecise <- module_run(paper, "imprecise_p")

imprecise$table # print table
```

You can use the `expand_text()` function to get the full sentence or paragraph for context. Here you can see that "p < .025" was not an imprecisely reported p-value, but a description of the preregistered alpha threshold. 

```{r}
exp <- expand_text(imprecise$table, paper)

exp$expanded # print expanded text
```

We can investigate the most common imprecise p-values in the PsychSci set. "p < .01" and "p < .05" are probably often describing figures or tables, but what is the deal with "p > .25"?

```{r}
imprecise_ps <- module_run(psychsci, "imprecise_p")

imprecise_ps$table |>
  count(text, sort = TRUE) |>
  head()
```

We can expand the text to check the context for "p > .25".

```{r}
gt.25 <- imprecise_ps$table |>
  filter(grepl("\\.25", text)) |>
  expand_text(paper = psychsci)

gt.25$expanded[1:3] # look at the first 3
```


### llm_summarise

Generate a 1-sentence summary for each section. This uses the `llm()` function with the query "summarise this section briefly, in one sentence" and returns a one-sentence summary of each major section (abstract, intro, method, results, discussion, acknowledgements, appendices)

```{r, eval = FALSE}
summaries <- module_run(paper, "llm_summarise", seed = 8675309)

summaries
```


|text |section |answer | time| tokens|
|:----------------------------------------------------------------------------------------------------|:----------|:----------------------------------------------------------------------------------------------------|---------:|------:|
|Across taxa, animals possess a number of behavioral adaptations that function to mitigate the cos... |abstract   |Humans have evolved behavioral adaptations to detect and avoid infectious microorganisms, includi... | 0.1327031|    373|
|the evolutionary-biology literature, which highlights the fact that organisms neutralize pathogen... |intro      |The study aimed to replicate a previous finding that people who have recently been ill are more a... | 0.2404155|   1406|
|Method Except where noted, all methodological details-including all stimuli and dot-probe procedu... |method     |The study replicated the methodological details of Miller and Maner (2011) to test the interactio... | 0.3714075|   2314|
|Results On the basis of our preregistered exclusion criteria, we removed the 9 participants with ... |results    |The study found a main effect of face type, with slower responses to disfigured faces, but no mai... | 0.2508097|   1227|
|Discussion The study replicated here has been interpreted as a key piece of evidence supporting a... |discussion |This study's null findings do not provide strong evidence that pathogen avoidance does not vary a... | 0.2339141|    904|
|NA Funding This work was supported by Horizon 2020 European Research Council Grant No. StG-2015 6... |funding    |This study was funded by a Horizon 2020 European Research Council Grant and has made its data, an... | 0.1644160|    210|
|Declaration of Conflicting Interests The author(s) declared that there were no conflicts of inter... |annex      |The author(s) declared that there were no conflicts of interest with respect to the authorship or... | 0.0778079|    101|

Showing 7 of 7 rows

The default table isn't very helpful here, so let's customise the output. The object returned from a module usually contains a table. We just want to show the section and answer columns here. 

```{r, echo = FALSE}
#saveRDS(summaries, "summaries.Rds")
summaries <- readRDS("summaries.Rds")
```



```{r}
summaries$table[, c("section", "answer")]
```

If you try to run this on the whole PsychSci set, you will get an error message because the number of LLM calls is much higher than the maximum number, which is set by default to 30 (the per-minute rate limit for the default model on a free groq account). You can increase the maximum allowable calls (`llm()` manages your rate limit, so this would take nearly an hour at 30 calls per minute) using `llm_max_calls()`, but we aren't going to do this here. 

```{r, error = TRUE}
summaries_ps <- module_run(psychsci, "llm_summarise", seed = 8675309)
```

### marginal

List all sentences that describe an effect as 'marginally significant'.

```{r, results='asis'}
marginal <- module_run(paper, "marginal")

marginal # print table
```

Let's check how many are in the full set.

```{r}
marginal_ps <- module_run(psychsci, "marginal")

marginal_ps$table # print table
```

### osf_check

List all OSF links and whether they are open, closed, or do not exist. This requires an internet connection to check the status of the link.

```{r}
osf_check <- module_run(paper, "osf_check")

osf_check$table
```

This would take a while to run for the 341 unique OSF links found in the PsychSci set (usually 1-2 seconds per link), so we will skip it here (but can reveal we found 10 closed links).

```{r, eval = FALSE}
osf_check_ps <- module_run(psychsci, "osf_check")
```

### ref_consistency

Check if all references are cited and all citations are referenced.

```{r}
ref_consistency <- module_run(paper, "ref_consistency")

ref_consistency$table
```

It looks like there are some references with missing citations. The first one doesn't look like a reference, and grobid often parses tables oddly. You'd need to look at the original PDF to see if the others are actually missing or false positives. Here, they are all false positives, based on grobid not being able to match the in-text citation to the reference list. 


### retractionwatch

Flag any cited papers in the RetractionWatch database. It takes a substantial time to make calls to crossref for each reference, so we provide a summary of the database in papercheck, which is updated regularly (last 2025-02-28). 

```{r}
rw <- module_run(paper, "retractionwatch")

rw$table
```

In the full PsychSci set, we find three papers in the RetractionWatch database cited, one with a retraction and two with corrections.  

```{r}
rw_ps <- module_run(psychsci, "retractionwatch")

rw_ps$table
```

Check the context of the citations in the text column:

```{r}
rw_ps$table$text
```

### statcheck

Check consistency of p-values and test statistics using functions from [statcheck](https://github.com/MicheleNuijten/statcheck).

```{r}
statcheck <- module_run(paper, "statcheck")

statcheck$table
```

Here we see a false positive, where the paper reported the results of an equivalence test, which are meant to be one-tailed, but statcheck did not detect that this was one-tailed.


In the full PsychSci set, there are more than 27K sentences with numbers to check, so this takes about a minute to run. 

```{r, results='asis', eval = FALSE}
statcheck_ps <- module_run(psychsci, "statcheck")
```

```{r, echo = FALSE}
#saveRDS(statcheck_ps, "statcheck_ps.Rds")
statcheck_ps <- readRDS("statcheck_ps.Rds")
```

There will be, of course, some false positives in the full set of `r nrow(statcheck_ps$table)` flagged values. Let's look just at the flagged values where the computed p-value is about double the reported p-value, and this changes the significance decision (at an alpha of 0.05).

```{r}
statcheck_ps$table |>
  filter(decision_error, 
         round(computed_p/reported_p, 1) == 2.0) |>
  select(reported_p, computed_p, raw) |>
  mutate(computed_p = round(computed_p, 4))
```

## Chaining Modules

Modules return a `summary` table as well as the detailed results `table`, which is automatically added to the summary if you chain modules.

```{r}
ps_metascience <- psychsci[1:10] |>
  module_run("all_p_values") |>
  module_run("imprecise_p") |>
  module_run("marginal")

ps_metascience$summary
```



