---
title: "Checking Preregistrations: prereg_check and reg_check"
output: rmarkdown::html_vignette
df-print: paged
vignette: >
  %\VignetteIndexEntry{Checking Preregistrations: prereg_check and reg_check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  eval = FALSE
)
```

```{r, eval = TRUE, include = FALSE}
library(metacheck)
```

## Why check preregistrations?

When researchers *preregister* a study, they write down their plan — how
many participants they will collect, what they predict, how they will
analyse the data — **before** they collect the data. This makes it possible
to distinguish between analyses that were planned in advance (confirmatory)
and analyses that emerged when looking at the data (exploratory).

But a preregistration only helps if readers actually compare it against the
published paper. Meta-scientific research shows this rarely happens:
deviations from preregistrations are common, often undisclosed, and most
frequently concern the sample size (van den Akker et al., 2024). Checking a
preregistration by hand is real work — you have to find it, read it, and
compare it line by line with the paper.

Metacheck offers two modules that make this work easier:

* **`prereg_check`** finds preregistration links in a paper, retrieves the
  preregistration, and organizes its content into a standard structure.
* **`reg_check`** goes one step further: it sends the preregistration and
  the paper to [RegCheck](https://github.com/JamieCummins/regcheck), a tool
  by Jamie Cummins that uses a large language model (LLM) to compare the two
  documents dimension by dimension, and flags where the paper might deviate
  from the plan.

**Important:** the LLM judgements from `reg_check` are *pointers for your
own checking*, not verdicts. Metacheck never automates quality evaluation —
a human always needs to verify the flagged deviations against the actual
documents.

## Step 1: Retrieve a preregistration with prereg_check

The `psychsci` dataset that comes with metacheck contains 250 open-access
papers from *Psychological Science*. Several of them link to
preregistrations on AsPredicted or the OSF. We use one of them here: a paper
that links to one AsPredicted preregistration.

```{r}
paper <- psychsci[["09567976231223130"]]

prereg <- module_run(paper, "prereg_check")
prereg
```

```
#> We found 1 preregistration.
```

The module found one preregistration (AsPredicted `sn9xs`), retrieved it,
and organized its content into named fields. The most useful ones for a
quick check:

```{r}
prereg$table[, c("id", "template_name", "sample_size")]
```

```{r, eval = TRUE, echo = FALSE}
data.frame(
  id = "sn9xs",
  template_name = "AsPredicted",
  sample_size = "We currently aim for a sample size of 30 participants. If more than 30 participants are collected, all collected participants will be included in our analyses."
) |> knitr::kable()
```

The full preregistration text is in `prereg$table` (one column per field of
the preregistration template: hypotheses, dependent variables, analyses,
exclusions, and so on). The module report shows all of this with links, so
you can verify everything at the source.

At this point you could compare the preregistration with the paper by hand
— and for any serious assessment, you should. But a first automated pass
can point you to the places worth looking at. That is what `reg_check`
does.

## Step 2: Compare paper and preregistration with reg_check

RegCheck compares the two documents on a set of *dimensions* — sample size,
hypotheses, exclusion criteria, statistical models, and so on. For each
dimension, an LLM reads the relevant passages from both documents and
judges:

* `"no"` — no deviation: paper and preregistration are consistent
* `"yes"` — potential deviation: the documents seem to differ
* `"missing"` — not enough information to judge

You can chain the two modules, so the preregistration is only retrieved
once:

```{r}
check <- paper |>
  module_run("prereg_check") |>
  module_run("reg_check", client = "groq")
```

This takes a while (about a minute with the hosted server, longer locally)
because the LLM makes one comparison per dimension. Here is the real result
for our example paper:

```{r}
check$table[, c("dimension", "deviation_judgement")]
```

```{r, eval = TRUE, echo = FALSE}
data.frame(
  dimension = c("Sample size", "Data source", "Inclusion criteria",
                "Exclusion criteria", "Incomplete and missing data",
                "Hypotheses", "Manipulated variables", "Measured variables",
                "Transformations", "Statistical models"),
  deviation_judgement = c("no", "yes", "yes", "no", "no", "no", "no", "no",
                          "missing", "yes")
) |> knitr::kable()
```

```
#> RegCheck compared the paper with 1 preregistration on 10 dimensions,
#> and flagged 3 potential deviations (1 dimension could not be assessed).
```

### Reading the results — and why you must verify them

For every dimension, `check$table` also contains a written explanation
(`deviation_information`), summaries of what each document says, and the
exact quotes the judgement is based on. The module report presents all of
this, with the evidence in a collapsible section.

Look at the three flagged deviations in our example:

* **Statistical models** (`yes`): the paper reports repeated-measures
  ANOVAs with factor combinations that the preregistration did not fully
  specify. That is exactly the kind of deviation worth checking — and worth
  reporting transparently (Lakens, 2024).
* **Data source** and **Inclusion criteria** (`yes`): RegCheck notes that
  the *paper* mentions undergraduate students and Mechanical Turk, while
  the *preregistration does not mention the data source at all*. Is that a
  "deviation"? Arguably it is an *omission in the preregistration* rather
  than a changed plan. This is a perfect example of why the judgements are
  pointers, not conclusions: the LLM flags a difference, and the human
  decides what it means.
* **Transformations** (`missing`): the preregistration says nothing about
  data transformations, so there is nothing to compare.

## Two ways to run RegCheck

### Option A: the hosted RegCheck server

The simplest option is the hosted RegCheck app (`client = "groq"` or
`"openai"`), which `reg_check` uses automatically when you pick one of
those clients. In our test, a full 10-dimension comparison took about 40
seconds.

The hosted server requires an **API token** in your `.Renviron` file (use
`usethis::edit_r_environ()` to open it):

```
REGCHECK_API_TOKEN="your_token_here"
```

Tokens are not publicly available: you currently need to request one from
the RegCheck team. Note that the hosted option sends the full paper and
preregistration text to the RegCheck server and the LLM provider behind it
(e.g., Groq) — do not use it for confidential manuscripts.

### Option B: a local RegCheck server with Ollama

A [fork of RegCheck](https://github.com/Matilda03/regcheck-fork) can run
the comparison entirely on your own computer, using
[Ollama](https://ollama.com) — the same free tool that metacheck uses for
other local LLM tasks (see the
[Local AI with Ollama](ollama.html) article). With this option, **no text
leaves your machine**, which makes it suitable for confidential work and
GDPR-sensitive contexts.

The local option is the default in `reg_check`:

```{r}
check <- paper |>
  module_run("prereg_check") |>
  module_run("reg_check")   # client = "ollama" is the default
```

Setting up the local server (once):

1. Install [Ollama](https://ollama.com/download) and pull a model, e.g.
   `ollama pull llama3.1:8b` in a terminal.
2. Clone the fork: `git clone https://github.com/Matilda03/regcheck-fork`
3. Follow its README to install the Python dependencies and start the
   server (it runs at `http://localhost:8000`).
4. Choose any token, configure it on the server as `REGCHECK_API_TOKEN`,
   and put the same token in your `.Renviron`. Yes — the local server also
   requires a token; it just stays on your machine.

### Is a small local model good enough?

The hosted server uses large commercial models. A local model that fits on
a normal laptop or desktop GPU is much smaller — is it up to the task?

We probed this with the largest model installed on our test machine
(`qwen3.5:9b`, 6.6 GB), giving it the same comparison task RegCheck uses
(same prompt structure, same judgement format) for three dimensions of our
example paper. The result:

* It produced valid, structured judgements on every call.
* Its judgements **agreed with the hosted model** on all three probed
  dimensions, with accurate, dimension-specific explanations (for example,
  it correctly identified the "first 30 non-excluded participants" rule in
  the sample size comparison).
* It took 25–55 seconds per dimension on a 6 GB GPU, so a full
  10-dimension comparison takes roughly 5–9 minutes per preregistration —
  compared to under a minute for the hosted server.

Two caveats. First, small models have a limited *context window*: they can
only read a limited amount of text at once. In our probe, feeding the whole
paper at once silently broke the comparison — the model must be given short
relevant excerpts, which is exactly what the RegCheck pipeline's retrieval
step does. Second, our probe covered three clear-cut dimensions; subtler
judgements (like the statistical-models deviation above) may favour larger
models. Our overall assessment: a ~8–9B local model is adequate for a first
screening pass, and the hosted option is preferable when you need the most
careful judgements and the manuscript is not confidential.

## Responsible use

* RegCheck judgements are **LLM output**. Verify every flagged deviation
  against the quoted evidence and the original documents before drawing
  conclusions — and expect some flags to be omissions or ambiguities rather
  than true deviations.
* Deviating from a preregistration is **not misconduct** — it is often
  justified or even necessary. The problem is *undisclosed* deviation. See
  Lakens (2024) for how to deviate well.
* For the empirical background on preregistration–paper consistency, see
  van den Akker et al. (2024).

## References

* van den Akker, O. R., et al. (2024). The potential of preregistration in
  psychology: Assessing preregistration producibility and
  preregistration-study consistency. *Psychological Methods*.
  doi: 10.1037/met0000687
* Lakens, D. (2024). When and How to Deviate From a Preregistration.
  *Collabra: Psychology*, 10(1), 117094. doi: 10.1525/collabra.117094
* RegCheck: Cummins, J., <https://github.com/JamieCummins/regcheck>;
  local Ollama fork: <https://github.com/Matilda03/regcheck-fork>
