---
title: "Checking Preregistrations: prereg_check and reg_check"
output: rmarkdown::html_vignette
df-print: paged
vignette: >
  %\VignetteIndexEntry{Checking Preregistrations: prereg_check and reg_check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  eval = FALSE
)
```

```{r, eval = TRUE, include = FALSE}
library(metacheck)
load(system.file("extdata", "regcheck_examples.rda", package = "metacheck"))
```

## Why check preregistrations?

When researchers *preregister* a study, they write down their plan — how many participants they will collect, what they predict, how they will analyse the data — **before** they collect the data. 
This makes it possible to distinguish between analyses that were planned in advance (confirmatory) and analyses that emerged when looking at the data (exploratory). 
Preregistration - when done well - has many benefits (Lakens, Mesquida, Rasti, & Ditroilo, 2024).  

But a preregistration only helps if readers actually compare it against the published paper. 
Meta-scientific research shows this rarely happens: deviations from preregistrations are common, often undisclosed, and most
frequently concern the sample size (van den Akker et al., 2024). 
Checking a preregistration takes time. 

Metacheck offers two modules that make this work easier:

* **`prereg_check`** finds preregistration links in a paper, retrieves the preregistration, and organizes its content into a standard structure.
* **`reg_check`** sends the preregistration and the paper to [RegCheck](https://github.com/JamieCummins/regcheck), a tool
  by Jamie Cummins that uses a large language model (LLM) to compare the two documents dimension by dimension, and flags where the paper might deviate from the plan.

**Important:** the LLM judgements from `reg_check` are *suggestions to consider*, not verdicts. Metacheck never automates quality evaluation — a human always needs to verify the flagged deviations against the actual documents. When Regcheck does a good job, it can assist researchers in spotting deviations more easily - but it is still the responsibility of a researcher to verify these deviations are actually present. 

## Example 1: AsPredicted preregistration

The `psychsci` dataset contains 250 open-access papers from *Psychological Science*. We start with a paper that links to one AsPredicted preregistration (`sn9xs`).

### Step 1: Retrieve the preregistration

```{r}
paper_ap <- psychsci[["09567976231223130"]]

prereg <- module_run(paper_ap, "prereg_check")
prereg
```

```
#> We found 1 preregistration.
```

```{r}
prereg$table[, c("id", "template_name", "sample_size")]
```

```{r, eval = TRUE, echo = FALSE}
data.frame(
  id = "sn9xs",
  template_name = "AsPredicted",
  sample_size = "We currently aim for a sample size of 30 participants. If more than 30 participants are collected, all collected participants will be included in our analyses."
) |> knitr::kable()
```

### Step 2: Compare with reg_check (hosted server)

Chain the two modules so the preregistration is retrieved once:

```{r}
check_ap <- paper_ap |>
  module_run("prereg_check") |>
  module_run("reg_check", client = "openai")
```

The `client` argument tells the hosted RegCheck server which LLM to use
for the comparison. The server supports three cloud backends:

| `client`     | Model used                  |
|:-------------|:----------------------------|
| `"openai"`   | GPT-5 (via OpenAI)          |
| `"groq"`     | Llama 3.3 70B (via Groq)    |
| `"deepseek"` | DeepSeek Reasoner           |

All three options send text to the hosted RegCheck server; the only
difference is which LLM processes it. We use `"openai"` here to get the
most capable model. The default client is `"ollama"`, which skips the
hosted server entirely and runs locally instead — covered in Example 3
below.

This takes roughly a minute on the hosted server. Here are the real results:

```{r}
check_ap$table[, c("dimension", "deviation_judgement")]
```

```{r, eval = TRUE, echo = FALSE}
check_ap$table[, c("dimension", "deviation_judgement")] |>
  knitr::kable()
```

```{r, eval = TRUE, echo = FALSE}
cat(check_ap$summary_text)
```

RegCheck flagged **1 potential deviation** (Statistical models) and found
3 dimensions with insufficient information to judge. The statistical models
dimension is a meaningful flag: the paper reports repeated-measures ANOVAs
with factor combinations that the preregistration did not fully specify.
The missing-information judgements for Data source, Inclusion criteria, and
Transformations reflect *omissions in the preregistration* rather than changed
plans — the LLM flags a gap, and the human decides what it means.

For every dimension, `check_ap$table` also contains `deviation_information`,
`paper_summary`, `prereg_summary`, `paper_quotes`, and `prereg_quotes` — the
evidence the LLM used. Always check these before drawing conclusions.

## Example 2: OSF preregistration

Now a paper that links to an OSF preregistration (`7qcxa`, using the
"OSF Preregistration" template):

```{r}
paper_osf <- psychsci[["0956797620904990"]]

check_osf <- paper_osf |>
  module_run("prereg_check") |>
  module_run("reg_check", client = "openai")
```

```{r}
check_osf$table[, c("dimension", "deviation_judgement")]
```

```{r, eval = TRUE, echo = FALSE}
check_osf$table[, c("dimension", "deviation_judgement")] |>
  knitr::kable()
```

```{r, eval = TRUE, echo = FALSE}
cat(check_osf$summary_text)
```

RegCheck flagged **5 potential deviations** for this paper, including sample
size, hypotheses, measured variables, transformations, and statistical models.
As always, verify each flag against the quoted evidence and the original
documents — some flags may reflect ambiguous wording in the preregistration
rather than true deviations.

## Generating a report

Once you have run `reg_check`, you can generate a standalone HTML report
with just the RegCheck section. Pass the paper to `report()` and specify
only the modules you want:

```{r}
paper <- psychsci[["09567976231223130"]]

report(paper,
       modules     = c("prereg_check", "reg_check"),
       args        = list(reg_check = list(client = "openai")),
       output_file = "regcheck_report_openai.html")
```

`report()` opens the report in your browser. Include `"prereg_check"` in
the module list so that `reg_check` can reuse the retrieved preregistration
without fetching it twice.

## Example 3: Running locally with Ollama

The same comparison can run entirely on your own machine. Metacheck bundles
a minimal local RegCheck server (inspired by the Ollama integration work by
[Matilda Gustafsson](https://github.com/Matilda03)) together with
[Ollama](https://ollama.com). No text leaves your machine, which makes it
suitable for confidential work and GDPR-sensitive contexts.

```{r}
check_local <- paper_ap |>
  module_run("prereg_check") |>
  module_run("reg_check")   # client = "ollama" is the default
```

Here are the real results from a local run on the same AsPredicted paper,
using `qwen3.5:9b` (a ~6 GB model):

```{r}
check_local$table[, c("dimension", "deviation_judgement")]
```

```{r, eval = TRUE, echo = FALSE}
check_local$table[, c("dimension", "deviation_judgement")] |>
  knitr::kable()
```

```{r, eval = TRUE, echo = FALSE}
cat(check_local$summary_text)
```

Several things to note when comparing the local and hosted results:

* **Different dimension labels.** The local model derived its own dimension
  names from the preregistration content, rather than using the standard
  RegCheck set (sample size, hypotheses, exclusion criteria, …). The standard
  set is defined server-side; small local models may ignore or rephrase it.
* **Blank judgements.** Several rows have an empty `deviation_judgement`.
  This happens when the local model does not follow the structured output
  format exactly — a known limitation of smaller models.
* **Where agreement holds.** For the dimensions where the local model did
  produce a valid judgement, it agreed with the hosted result on flagging
  statistical analysis / statistical models as a deviation (`yes`).

Local runs took 25–90 seconds per dimension on a 6 GB GPU, so a full
10-dimension comparison takes roughly 5–15 minutes — compared to under a
minute for the hosted server.

To generate a report from a local run, use the same `report()` call with
`client = "ollama"` (the default) and start the local server first:

```{r}
paper <- demopaper()

regcheck_start_local()

report(paper,
       modules     = c("prereg_check", "reg_check"),
       output_file = "regcheck_report_ollama.html")
```

## Two ways to run RegCheck

### Option A: the hosted RegCheck server

Use `client = "groq"`, `"openai"`, or `"deepseek"`. The hosted server
requires an **API token** in your `.Renviron` file:

```
REGCHECK_API_TOKEN="your_token_here"
```

Use `usethis::edit_r_environ()` to open the file. Tokens are not publicly
available: contact the RegCheck team. Note that the hosted option sends the
full paper and preregistration text to the RegCheck server and the
corresponding LLM provider — do not use it for confidential manuscripts.

### Option B: a local RegCheck server with Ollama

The RegCheck server is bundled inside metacheck, so there is nothing to
clone or configure separately. You need two external tools installed once:
[Ollama](https://ollama.com/download) and either
[Docker](https://www.docker.com/get-started/) (easiest) or
[Python 3.12+](https://www.python.org/downloads/).

No API token configuration is needed — metacheck uses a built-in default
token for local use automatically.

#### Option B1: Docker (recommended)

Docker gives you a fully isolated, reproducible environment with no Python
installation needed beyond what Docker manages.

**Step 1: Install Docker**

Download Docker Desktop from <https://www.docker.com/get-started/> and
start it. Verify the install by running `docker --version` in a terminal.

**Step 2: Install Ollama and pull the models**

Download from <https://ollama.com/download>. Then pull the embedding model
(required) and a language model of your choice:

```bash
ollama pull nomic-embed-text-v2-moe
ollama pull llama3.2
```

**Which language model to use?** Larger models produce better judgements —
use the largest one that fits in your GPU memory. A few options from the
[Ollama library](https://ollama.com/library):

| Model               | Size   | Minimum VRAM |
|:--------------------|:-------|:-------------|
| `llama3.2`          | 3B     | ~4 GB        |
| `mistral`           | 7B     | ~8 GB        |
| `llama3.1`          | 8B     | ~8 GB        |
| `llama3.3`          | 70B    | ~48 GB (or run on CPU, slowly) |

Replace `llama3.2` with your chosen model name. To tell the local server
which model to use, pass it when starting (see Step 3).

The embedding model (`nomic-embed-text-v2-moe`) is required regardless of
which language model you pick.

**Step 3: Start the server from R**

```{r}
regcheck_start_local()                         # uses llama3.2 by default
regcheck_start_local(model = "llama3.1")       # use a different model
```

The first run builds the Docker image (roughly 5 minutes, once only).
After that, `regcheck_start_local()` is instant. The server runs at
`http://localhost:8000`. Stop it when you are done:

```{r}
regcheck_stop_local()
```

Then run the comparison as usual (the default `client = "ollama"` uses
your local server):

```{r}
paper <- demopaper()
check <- paper |>
  module_run("prereg_check") |>
  module_run("reg_check")
```

#### Option B2: Manual Python (no Docker)

If you prefer not to install Docker, you can run the server directly with
Python.

**Step 1: Install Python 3.12+**

Download from <https://www.python.org/downloads/> and make sure `python3`
(or `python` on Windows) is on your PATH.

**Step 2: Install Ollama and pull models**

Download from <https://ollama.com/download>. Pull the embedding model and
your chosen language model (see the table in Option B1 above for guidance
on which model to pick):

```bash
ollama pull nomic-embed-text-v2-moe
ollama pull llama3.2          # or any larger model you prefer
```

**Step 3: Set up the server from R (once)**

```{r}
regcheck_setup_local()
```

This creates a Python virtual environment inside the metacheck package
directory, installs all dependencies (no Redis required — the bundled
server uses an in-memory store), and downloads the NLTK data.

**Step 4: Start the server**

Each R session, before running `reg_check`:

```{r}
regcheck_start_local(method = "python")
```

The server runs at `http://localhost:8000`. Stop it when you are done:

```{r}
regcheck_stop_local()
```

Then run the comparison:

```{r}
paper <- demopaper()
check <- paper |>
  module_run("prereg_check") |>
  module_run("reg_check")
```

### Is a small local model good enough?

The hosted server uses large commercial models. A local model that fits on
a typical laptop or desktop GPU is much smaller — is it up to the task?

Based on our real-world test above (Example 3):

* A ~8–9B model produces valid judgements on most dimensions, and its
  judgements can agree with the hosted model on clear-cut cases.
* Small models frequently do not respect the structured output format,
  leaving `deviation_judgement` blank for some rows.
* Small models tend to derive their own dimension labels rather than
  following the server's standard set, making results harder to compare
  across papers.
* A full comparison takes 5–15 minutes locally vs. under a minute hosted.

**Our overall assessment:** a local model is adequate for a first
exploratory pass on confidential manuscripts. For systematic screening
or the most careful judgements, the hosted option is preferable.

## Responsible use

* RegCheck judgements are **LLM output**. Verify every flagged deviation
  against the quoted evidence and the original documents before drawing
  conclusions — and expect some flags to be omissions or ambiguities rather
  than true deviations.
* Deviating from a preregistration is **not misconduct** — it is often
  justified or even necessary. The problem is *undisclosed* deviation. See
  Lakens (2024) for how to deviate well.
* For the empirical background on preregistration–paper consistency, see
  van den Akker et al. (2024).

## References

* van den Akker, O. R., et al. (2024). The potential of preregistration in
  psychology: Assessing preregistration producibility and
  preregistration-study consistency. *Psychological Methods*.
  doi: 10.1037/met0000687
* Lakens, D., Mesquida, C., Rasti, S., & Ditroilo, M. (2024). The benefits of preregistration and Registered Reports.           Evidence-Based Toxicology, 2(1). https://doi.org/10.1080/2833373X.2024.2376046
* Lakens, D. (2024). When and How to Deviate From a Preregistration.
  *Collabra: Psychology*, 10(1), 117094. doi: 10.1525/collabra.117094
* RegCheck: Cummins, J., <https://github.com/JamieCummins/regcheck>
* Local Ollama integration (inspiration): Gustafsson, M., <https://github.com/Matilda03>
