---
title: "Text Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Text Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| message: false

library(papercheck)
library(dplyr)

set.seed(8675309) 
```


You can build a simple general linear model to predict the classification of text strings.

## Set up ground truth

See the [metascience vignette](metascience.html) for an explanation of how to set up a ground truth table. Here, we're going to split our data into a training and test set.

```{r}
ground_truth <- readxl::read_excel("power/power_screening_coded.xlsx")

train <- dplyr::slice_sample(ground_truth, prop = 0.5)
test <- anti_join(ground_truth, train, by = "text")
```




## Get important words

You can use any method for finding the words you want to use in your model, but papercheck has a built-in function to find the words that are most distinctive in your classification groups. The classification values here are 0 and 1, but can be TRUE/FALSE or any two text values.

`n_X` is the total number of incidents of the word in category X, while `freq_X` is the average number of incidents per text string in category X (so can be higher than 1 if a word tends to be found several times per sentence). The table gives you the top `n` words with the largest absolute difference in frequency.

```{r}
words <- distinctive_words(
  text = train$text,
  classification = train$power_computation,
  n = 10
)
```

```{r, echo = FALSE}
knitr::kable(words, digits = 2)
```


By default, the function will "stem" words using the "porter" algorithm. For example, "sampl" will match "sample", "samples" and "sampling". If your text is not English, check `SnowballC::getStemLanguages()` for other supported languages, or set `stem_language = FALSE`. 

You can get rid of words that you think will be irrelevant (even if they are predictive of classification in this data set) by adding them to `stop_words`. The `tidytext::stop_words` object gives you a list of common stop words, but this includes words like "above", "according", or "small", so use this with caution. 

The "###" value represents any number (the default setting for the `numbers` argument). We can set the `numbers` argument to "specific" to see if there are any specific numbers associated with power analyses.

```{r}
words <- distinctive_words(
  text = train$text,
  classification = train$power_computation,
  n = 10,
  numbers = "specific",
  stop_words = c("the", "a", "of", "an", "and")
)
```

```{r, echo = FALSE}
knitr::kable(words, digits = 2)
```

## Code text features

Next, code the features of your ground truth text using `text_features()`. This will give you a data frame that codes 0 or 1 for the absence or presence of each word or feature.

* `word_count` defaults to TRUE, and returns the number of words in each text string. 
* `has_number` defaults to TRUE, and checks for any number in your text. If "###" is in your words list, this will be automatically set to TRUE. 
* `has_symbols` is a named vector of non-word strings (use regex) that you want to detect. 
* `values` defaults to "presence" and returns 0 or 1 for the presence of a word in each text string, while "count" returns the number of incidences of the word per string. 

```{r}
has_symbols <- c(has_equals = "=", 
                 has_percent = "%")

features <- text_features(
  text = train$text, 
  words = words,
  word_count = FALSE, 
  has_number = TRUE,
  has_symbol = has_symbols, 
  values = "presence" # presence or count
)

# show the first row
features[1, ] |> str()
```

## Train a model

You can then use this feature data to train a model. Here, we're using a simple binomial logistic regression to predict the classification from all of the features.

```{r}
# Train logistic regression model
model <- glm(train$power_computation ~ .,
             data = features,
             family = "binomial")

summary(model)
```

You can use any model you like and any method to assess and choose the best model.


## Predict classification

Now you can classify any text using this model. First, we're going to predict the classification of the original training data. Use `text_features()` to get the feature data and `predict()` to return the model response, and compare this result to a threshold (here 0.5) to generate the predicted classification.

```{r}
train$model_response <- predict(model, features)

train$power_computation_predict <-
  train$model_response > 0.5

dplyr::count(train, 
             power_computation, 
             power_computation_predict)
```

Now you should test this on a new set of data.

```{r}

test_features <- text_features(
  text = test$text, 
  words = words,
  word_count = FALSE, 
  has_number = TRUE,
  has_symbol = has_symbols, 
  values = "presence" # presence or count
)
test$model_response <- predict(model, test_features)

test$power_computation_predict <-
  test$model_response > 0.5

dplyr::count(test, 
             power_computation, 
             power_computation_predict)
```


<!--

### Create a module

To create a module for this, you will need to save your model and create a module with code.

```{r, eval = FALSE}
saveRDS(model, "power/power_log_model.Rds")
```


```
{
  "title": "Classify with Log Model",
  "description": "",
  "code": {
    "packages": ["papercheck"],
    "code": "power_log_model.R"
  }
}
```


```{r, eval = FALSE}

words <- c("size", "effect", "sample", 
           "participants", "detect", "analysis",
           "medium", "power", "faul", "respondents", 
           "α", "cohen's", "research", 
           "version", "experiment")

has_symbols <- c(has_equals = "=", 
                 has_percent = "%")

readRDS("power/power_log_model.Rds")

table <- paper |>
  papercheck::search_text("power(ed)?\\b") |>
  papercheck::search_text("(\\.[0-9]|[0-9]%)")

features <- papercheck::text_features(
  table$text, words, has_symbols = has_symbols
)

table$model_response <- predict(model, features)
table$model_classification <- table$model_response > 0.5

traffic_light <- "red"
if (any(table$model_classification)) traffic_light <- "green"

# return
list(
  table = table,
  traffic_light = traffic_light
)

```

```{r, eval = FALSE}
v4 <- module_run(psychsci, "power/power_log_model.mod")
```


-->



